### 2014-11-10 Abstract.tex
Pass up to end of Introduction

We present a new framework for modular verification of hardware designs in the
style of the Bluespec language.  That is, we formalize the idea of components
in a hardware design, with well-defined input and output channels; and we show
how to specify and verify components individually, with machine-checked proofs
in the Coq proof assistant.  As a demonstration, we verify a fairly realistic
implementation of a multicore shared-memory system with two types of
components:  memory-system and processor.  Both components include nontrivial
optimizations, with the memory system employing an arbitrary hierarchy of cache
nodes that communicate with each other concurrently, and with the processor
doing speculative execution of many concurrent read operations.  Nonetheless,
we prove that the combined system implements sequential consistency.  One
interesting aspect of the proof is that the memory component is well-specified
to be reusable for systems implementing a variety of weak memory models.  To
our knowledge, our memory-system system proof is the first machine verification
of a cache-coherence protocol parameterized over an arbitrary cache hierarchy,
and our full-system proof is the first machine verification of sequential
consistency for a multicore hardware design that includes caches and
speculative processors.

>>>>

We present a new framework for modular verification of hardware designs in the
style of the Bluespec language.  That is, we formalize the idea of components
in a hardware design, with well-defined input and output channels; and we show
how to specify and verify components individually, with machine-checked proofs
in the Coq proof assistant.  As a demonstration, we verify a fairly realistic
implementation of a multicore shared-memory system with two types of
components: memory system and processor.  Both components include nontrivial
optimizations, with the memory system employing an arbitrary hierarchy of cache
nodes that communicate with each other concurrently, and with the processor
doing speculative execution of many concurrent read operations.  Nonetheless,
we prove that the combined system implements sequential consistency.  One
interesting aspect of the proof is that the memory component is well-specified
to be reusable for systems implementing a variety of weak memory models.  To
our knowledge, our memory-system proof is the first machine verification
of a cache-coherence protocol parameterized over an arbitrary cache hierarchy,
and our full-system proof is the first machine verification of sequential
consistency for a multicore hardware design that includes caches and
speculative processors.

### 2014-11-10 Introduction.tex
Pass up to end of Introduction

Modularity has long been understood as a key property for effective design and
verification of such complex systems. For the designer, it allows a separation
of concerns, increasing robustness by allowing the behavior encapsulated by a
modular boundary to be realized by multiple implementations, any of which may
be dropped into the system safely. It also allows a greater parallelization of
human design effort, improving development time. Similarly, verification is
simplified, as modular interface agreements provide a natural lemma structure.
This leads us towards a decomposition of the whole system verification task
into lemmas of subsystems, which can be composed in a black-box manner to
produce full-system theorems.

@@

LTSes as hardware descriptions are an established
idea \cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are compilers that
convert LTSes into efficient hardware.  Our work is based on the Bluespec
language \cite{BSV:LangRef, Bluespec:TFRG}, which served as the model for the
formalism of this paper.  Bluespec models hardware components as atomic rules
of a transition system over state elements, and there is a commercial compiler
to synthesize such codes into register transfer-level circuits (i.e., Verilog code) with
competitive performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation \cite{DNA:CoherenceImplementation} used to implement a
cycle-accurate simulator for a multi-processor PowerPC system \cite{Khan:PowerPc}. The hardware synthesized
from that implementation is rather efficient: an 8-core system with a 2-level
cache hierarchy can run 55 million instructions per second on the BEE FPGA
board \cite{bee2}. Figure \ref{both} gives an example of an LTS inference rule
transliterated directly from an actual Bluespec rule used in the PowerPC
system; its meaning will become clear in Section \ref{sec:cc}. In Bluespec
notation, only those state elements whose values change are mentioned in the
rule; in contrast LTSes employ a more functional notation.

The main contributions of this paper are:
\vspace{-.3cm}
\begin{itemize} 
\item Identification of an interface between the processor and the memory
components of a multi-processor system that enables modular implementation and
verification.

@@

\paragraph{Paper Organization:} We begin with a discussion of related work in
Section~\ref{relatedWork}. In Section~\ref{sec:lts} we introduce our flavor of
the labeled transition systems formalism, including a definition of trace
refinement. In Section~\ref{sec:store-atomicity}, we show a generic
decomposition of any multi-processor system independently of the memory model
that it implements, and we discuss the store-atomicity property of the memory
sub-component. In Section~\ref{sec:sc} we give a simple formal model of
sequential consistency.  The following sections refine the two main
subcomponents of our multi-processor system.  Section~\ref{sec:ooo} discusses
definition and verification of a speculative processor model, and
Section~\ref{sec:cc} defines and proves our hierarchical cache-coherence
protocol.  Finally, in Section~\ref{sec:finalresult} we show the whole-system
modular proof of our complex system, and end with some conclusions in
Sections~\ref{sec:conclusion}.

\begin{figure}[t]
\centering
\begin{boxedminipage}{\columnwidth}
\small
\noindent \textbf{rule} $(c.\parent == p \; \mathit{\&\&} \; c.\s[a] < x \; \mathit{\&\&} \; c.\wt[a]==$None$);$

>>>>

Modularity has long been understood as a key property for effective design and
verification of such complex systems. For the designer, it allows a separation
of concerns, increasing robustness by allowing the behavior encapsulated by a
modular boundary to be realized by multiple implementations, any of which may
be dropped into the system safely. It also allows a greater parallelization of
human design effort, improving development time. Similarly, verification is
simplified, as modular interface agreements provide a natural lemma structure.
They lead us towards a decomposition of the whole-system verification task
into lemmas about subsystems, which can be composed in a black-box manner to
produce full-system theorems.

@@

LTSes as hardware descriptions are an established
idea \cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are compilers that
convert LTSes into efficient hardware.  Our work is based on the Bluespec
language \cite{BSV:LangRef, Bluespec:TFRG}, which served as the model for the
formalism of this paper.  Bluespec models hardware components as atomic rules
of a transition system over state elements, and there is a commercial compiler
to synthesize such codes into register-transfer-level circuits (i.e., Verilog code) with
competitive performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation \cite{DNA:CoherenceImplementation} used to implement a
cycle-accurate simulator for a multiprocessor PowerPC system \cite{Khan:PowerPc}. The hardware synthesized
from that implementation is rather efficient: an 8-core system with a 2-level
cache hierarchy can run 55 million instructions per second on the BEE FPGA
board \cite{bee2}. Figure \ref{both} gives an example of an LTS inference rule
transliterated directly from an actual Bluespec rule used in the PowerPC
system.  Its meaning will become clear in Section \ref{sec:cc};
for now we just note that, while Bluespec rules employ Boolean \emph{guard}
conditions paired with sequences of imperative \emph{actions} to update state,
our (equivalent) formalization follows a natural-deduction inference-rule style,
where we explain a state transition as a relation between two tuples of state
elements, using rule premises where Bluespec uses guards.  The high-level point
is that the natural-deduction style is a semantically trivial rephrasing of the
Bluespec notation, which has been demonstrated as practical for synthesizing
realistic, high-performance hardware.  Inference rules like the one in
Figure \ref{both} fit naturally with the inductive-definition mechanisms of
proof systems like Coq, providing an effective starting point for formal
verification of hardware designs.

The main contributions of this paper are:
\vspace{-.3cm}
\begin{itemize} 
\item Identification of an interface between the processor and the memory
components of a multiprocessor system that enables modular implementation and
verification.

@@

\paragraph{Paper Organization:} We begin with a discussion of related work in
Section~\ref{relatedWork}. In Section~\ref{sec:lts} we introduce our flavor of
the labeled transition systems formalism, including a definition of trace
refinement. In Section~\ref{sec:store-atomicity}, we show a generic
decomposition of any multiprocessor system independently of the memory model
that it implements, and we discuss the store-atomicity property of the memory
subcomponent. In Section~\ref{sec:sc} we give a simple formal model of
sequential consistency.  The following sections refine the two main
subcomponents of our multiprocessor system.  Section~\ref{sec:ooo} discusses
definition and verification of a speculative processor model, and
Section~\ref{sec:cc} defines and proves our hierarchical cache-coherence
protocol.  Finally, in Section~\ref{sec:finalresult} we show the whole-system
modular proof of our complex system, and we end with some conclusions in
Section~\ref{sec:conclusion}.

\begin{figure}[t]
\centering
\begin{boxedminipage}{\columnwidth}
\small
\noindent \textbf{rule} $(c.\parent == p \; \mathit{\&\&} \; c.\s[a] < x \; \mathit{\&\&} \; c.\wt[a]== \textrm{None});$
